\ttitle

### 2014-01-15 Chapters/Chapter4.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

\TODO{Decide on name}
\begin{itemize}
  \item Human Activity Support Vector Change Detection
  \item Inertial Sensor Time Series Svm Temporal Segmentation
  \item Support Vector bases Human Activity Temporal Segmentation
  \item Human Activity Temporal Segmentation by Support Vector Machines: HATS-SVM
  \item HATS with One-Class SVM: HATS-OCS, HATS-OS, HATSOS
  \item OCS-HATS
  \item HATS-SVDD
\end{itemize}

@@

This chapter will introduce our method and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

>>>>

@@

This chapter will introduce our method, which we named \gls{ocs-hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

### 2014-01-15 Chapters/Chapter4/change_detection.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
This will label the new data point as an in- or outlier, relative to the current data set.
Depending on that outcome, the model is updated to include the new data point.
The second step is to inspect the (changed) model properties, \ie the radius $R$.
In our method we only employ the second step: the inspection of the radius $R$.
We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boundary include the new data point $\vectorsym{z}$.
Would the new data object $\vectorsym{z}$ be already an inlier, then the updated model would not have a significant change in radius $R$.
Our algorithm is outlined in \Cref{tab:algorithm_proposed_method}.

>>>>

\begin{algorithm}
\caption{\acrlong{ocs-hats} algorithm.}
\label{alg:ocs-hats}
\begin{algorithmic}[1]
\Require time series $\vectorsym{x}$ and window size $n$
\Ensure collection of change points $\vectorsym{cp}$

\State $w \leftarrow \vectorsym{x}[1 \dots n-1]$ \Comment{Initialize working set}

\While{$\vectorsym{x}_t$ is available}
  \State $w[] \leftarrow \vectorsym{x}_t$ \Comment{Add new data to working set}

  \State $s \leftarrow \operatorname*{SVDD}(\vectorsym{x})$ \Comment{Create hypersphere with SVDD method}
  \State $r[] \leftarrow s.radius$ \Comment{Extract radius from hypersphere}

  \State $r_{avg} \leftarrow \operatorname*{average}(r)$
  \State $\hbar \leftarrow s.radius / r_{avg}$ \Comment{Radius ratio $\hbar$}

  \If{ $\hbar$ \textless $th_{low}$ \textbf{or} $\hbar$ \textgreater $th_{high}$} \Comment{Compare $\hbar$ with thresholds}
    \State $cp[] \leftarrow t$ \Comment{Mark $t$ as change point}
  \EndIf

  \State $w[0] \leftarrow []$ \Comment{Drop oldest data point from working set}
\EndWhile

\State $cp \leftarrow \operatorname*{merge}(cp)$ \Comment {Post-processing, change point merging}
\end{algorithmic}
\end{algorithm}

\TODO{Make algorithm more readable}

In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
This will label the new data point as an in- or outlier, relative to the current data set.
Depending on that outcome, the model is updated to include the new data point.
The second step is to inspect the (changed) model properties, \ie the radius $R$.
In \gls{ocs-hats} we only employ the second step: the inspection of the radius $R$.
We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boundary include the new data point $\vectorsym{z}$.
Would the new data object $\vectorsym{z}$ be already an inlier, then the updated model would not have a significant change in radius $R$.
Our algorithm is outlined in \Cref{tab:algorithm_proposed_method}.

### 2014-01-15 Chapters/Chapter4/data_gathering.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which our method is based.
These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
Over the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in \Cref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
For our setup we record the activities of humans performed both in- and outdoor in an unknown environment.
Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
Our method uses the signals from the accelerometer, magnetic field, and rotation sensors.
These time series data are used to detect change points.
A video recording from the performed activity is used to annotate the time series with real change points.
The discovered change points are compared with these annotated change points to give a subjective quality measure.
In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

>>>>

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which \gls{ocs-hats} is based.
These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
Over the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in \Cref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method for change detection and temporal segmentation to real-world data sets.
For our setup we record the activities of humans performed both in- and outdoor in an unknown environment.
Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
\gls{ocs-hats} uses the signals from the accelerometer, magnetic field, and rotation sensors.
These time series data are used to detect change points.
A video recording from the performed activity is used to annotate the time series with real change points.
The discovered change points are compared with these annotated change points to give a subjective quality measure.
In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

### 2014-01-15 Chapters/Chapter6/data_sets.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

% !TEX root = ../../main.tex
\section{Data Sets}\label{sec:data_sets}
In the previous chapter we have applied our method to the artificial data sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
In this chapter we apply the method to real-world data sets.
The data sets come from inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
To compare our results with other research, we have tried to apply our method to two commonly used data sets.
The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.
As an excerpt from the data in \Cref{fig:wisdm_excerpt} shows, the activities are recorded non-continuous.
Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
Even if the gap between the activities would be removed, the transition segments between activities would still be missing and thus the data set would not reflect continuous recordings.
The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
As shown in an except in \Cref{fig:uci_annotated}, the labeling to the data segments seems to be incorrect.
Furthermore, it is not clear how the data is recorded and whether all activities are performed continuously.

>>>>

% !TEX root = ../../main.tex
\section{Data Sets}\label{sec:data_sets}
In the previous chapter we have applied our method \gls{ocs-hats} to the artificial data sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
In this chapter we apply the method to real-world data sets.
The data sets come from inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
To compare our results with other research, we have tried to apply \gls{ocs-hats} to two commonly used data sets.
The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.
As an excerpt from the data in \Cref{fig:wisdm_excerpt} shows, the activities are recorded non-continuous.
Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
Even if the gap between the activities would be removed, the transition segments between activities would still be missing and thus the data set would not reflect continuous recordings.
The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
As shown in an except in \Cref{fig:uci_annotated}, the labeling to the data segments seems to be incorrect.
Furthermore, it is not clear how the data is recorded and whether all activities are performed continuously.

### 2014-01-15 Chapters/Chapter6/results.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}
In this section we will provide the results of our methods applied to the real-world data sets and discuss the quality of the solution.
As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} and $\operatorname*{Average\_delay}$ (following \Cref{sec:artificial_data_results}).
Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.

>>>>

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}
In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} and $\operatorname*{Average\_delay}$ (following \Cref{sec:artificial_data_results}).
Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.

### 2014-01-15 Chapters/Chapter7.tex
Finished feedback Chapter 1
Decide on OCS-HTS. Working on algorithm in chapter 4

In this research we have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
This was performed in a real-world setting, showing the robustness of the method for noisy data.
Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we have put emphasis on the subjective inspection of the results.
The proposed method gives fairly good results, although parameter tuning is a subject of furter research.

Our method is based on and a slightly simplification of the \gls{svcpd} by Camci~\cite{camci2010change}.
We have tested our proposed method with the same artificial data and did not find a deterioration in the results.

@@

Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
This allows for better interpretation of the results.
All the recorded data is made publicity available~\cite{vlasveld2013continuous}.

>>>>

In this research we have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
This was performed in a real-world setting, showing the robustness of the method for noisy data.
Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we have put emphasis on the subjective inspection of the results.
The proposed method gives fairly good results, although parameter tuning is a subject of further research.

We have constructed \gls{ocs-hats} as a slightly simplification of \gls{svcpd} by Camci~\cite{camci2010change} application to real-world inertial sensor data.
We have tested our proposed method with the same artificial data and did not find a deterioration in the results.

@@

Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
This allows for better interpretation of the results.
All the recorded data is made publicity available under the \gls{almende-data} data set \cite{vlasveld2014acras}.
